Approximation as Differentiation

### 2025-04-15 appendix/notes/cmon-enriched.tex
Clarify enriched functor properties.

\begin{definition}[$\mathscr{V}$-enriched functor]
For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $\cat{C}$ and
$\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F: \cat{C} \to \cat{D}$ on morphisms is a family of
morphisms $F_{X,Y}: \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$
that makes the following diagrams in $\mathscr{V}$ commute:

\begin{center}
\begin{tikzcd}[column sep=1.8cm]
I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{C}{X}{X} \arrow[d, "F_{X,X}"] \\
I \arrow[r, "\id_{F(X)}"'] & \Hom{D}{F(X)}{F(X)}
\end{tikzcd}
\hspace{5mm}
\begin{tikzcd}[column sep=2.6cm]
   \Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"']
   & \Hom{\cat{C}}{X}{Z} \arrow[d, "F_{X,Z}"] \\
   \Hom{\cat{D}}{F(Y)}{F(Z)} \tensor \Hom{\cat{D}}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] &
   \Hom{\cat{D}}{F(X)}{F(Z)}
\end{tikzcd}
\end{center}
\end{definition}

@@

In the specific setting of a $\CMon$-enriched presheaf $F: \cat{C}^\op \to \CMon$, enrichment means that $F$
preserves the additive structure of morphisms:
\begin{itemize}
\item $F(0_{X,Y}) = 0_{F(Y),F(X)}$
\item $F(f + g) = F(f) + F(g)$
\end{itemize}
It then follows automatically that $F$ also preserves bilinear composition, i.e.:
\begin{itemize}
\item $F(f) \comp F(0) = F(f) = F(0) \comp F(f)$
\item $F(f + g) \comp F(h) = F(h \comp f) + F(h \comp g)$
\item $F(h) \comp F(f + g) = F(f \comp h) + F(g \comp h)$
\end{itemize}
in addition to the usual enriched functor properties.

@@

Suppose $\cat{C}$ a small category. The usual \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda:
\cat{C} \to \Func{\cat{C}^\op}{\Set}$ sending:
\begin{itemize}
\item any object $X$ in $\cat{C}$ to the contravariant hom-functor $\Yoneda(X) = \Hom{\cat{C}}{-}{X}:
\cat{C}^\op \to \Set$
which sends:
   \begin{itemize}
   \item any object $Y$ in $\cat{C}$ to the hom-set $\Hom{\cat{C}}{Y}{X}$;
   \item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
   \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$.
   \end{itemize}
\item any morphism $f: X \to Y$ in $\cat{C}$ to the natural transformation $\Yoneda(f)$ where $\Yoneda(f)_Z =
(f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
\end{itemize}

\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the
hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is $\mathscr{V}$-enriched. The following
diagram commutes because precomposition with the identity is the identity on morphisms:

>>>>

\begin{definition}[$\mathscr{V}$-enriched functor]
\label{def:cmon-enriched:enriched-functor}
For any monoidal category $(\mathscr{V}, \tensor, I)$ and $\mathscr{V}$-enriched categories $\cat{C}$ and
$\cat{D}$, a \emph{$\mathscr{V}$-enriched} functor $F: \cat{C} \to \cat{D}$ on morphisms is a family of
morphisms $F_{X,Y}: \Hom{\cat{C}}{X}{Y} \to \Hom{\cat{D}}{F(X)}{F(Y)}$ between hom-objects in $\mathscr{V}$
that makes the following diagrams in $\mathscr{V}$ commute:

\begin{center}
\begin{tikzcd}[column sep=1.8cm]
I \arrow[d, equals] \arrow[r, "\id_X"] & \Hom{\cat{C}}{X}{X} \arrow[d, "F_{X,X}"] \\
I \arrow[r, "\id_{F(X)}"'] & \Hom{\cat{D}}{F(X)}{F(X)}
\end{tikzcd}
\hspace{5mm}
\begin{tikzcd}[column sep=2.6cm]
   \Hom{\cat{C}}{Y}{Z} \tensor \Hom{\cat{C}}{X}{Y} \arrow[r, "\comp_{X,Y,Z}"] \arrow[d, "F_{Y,Z} \tensor F_{X,Y}"']
   & \Hom{\cat{C}}{X}{Z} \arrow[d, "F_{X,Z}"] \\
   \Hom{\cat{D}}{F(Y)}{F(Z)} \tensor \Hom{\cat{D}}{F(X)}{F(Y)} \arrow[r, "\comp_{F(X),F(Y),F(Z)}"'] &
   \Hom{\cat{D}}{F(X)}{F(Z)}
\end{tikzcd}
\end{center}
\end{definition}

\noindent These generalise the usual functor laws to a setting where the hom-objects are not necessarily sets
and thus do not support an elementwise formulation.

@@

In the specific setting of a $\CMon$-enriched presheaf $F: \cat{C}^\op \to \CMon$, enrichment means that $F$
preserves the additive structure of morphisms:
\begin{itemize}
\item $F(0_{X,Y}) = 0_{F(Y),F(X)}$
\item $F(f + g) = F(f) + F(g)$
\end{itemize}
It then follows automatically that $F$ also preserves bilinear composition, i.e.:
\begin{itemize}
\item $F(f) \comp F(0) = F(f) = F(0) \comp F(f)$
\item $F(f + g) \comp F(h) = F(h \comp f) + F(h \comp g)$
\item $F(h) \comp F(f + g) = F(f \comp h) + F(g \comp h)$
\end{itemize}
in addition to the usual enriched functor properties in \defref{cmon-enriched:enriched-functor}.

@@

Suppose $\cat{C}$ a small category. The usual \emph{Yoneda embedding} for $\cat{C}$ is the functor $\Yoneda:
\cat{C} \to \Func{\cat{C}^\op}{\Set}$ defined as follows:
\begin{itemize}
\item on objects: $\Yoneda(X) = \Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \Set$, the contravariant hom-functor
sending:
   \begin{itemize}
   \item any object $Y$ in $\cat{C}$ to the hom-set $\Hom{\cat{C}}{Y}{X}$;
   \item any morphism $f: Y \to Z$ in $\cat{C}$ to the function $\Hom{\cat{C}}{f}{X} = (- \comp f):
   \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Y}{X}$.
   \end{itemize}
\item on morphisms: for any $f: X \to Y$ in $\cat{C}$, $\Yoneda(f)$ is the natural transformation where
$\Yoneda(f)_Z = (f \comp -): \Hom{\cat{C}}{Z}{X} \to \Hom{\cat{C}}{Z}{Y}$.
\end{itemize}

\noindent Now suppose $\cat{C}$ is $\mathscr{V}$-enriched with $\mathscr{V}$ closed monoidal; then the
hom-functor $\Hom{\cat{C}}{-}{X}: \cat{C}^\op \to \mathscr{V}$ is itself $\mathscr{V}$-enriched. The

The following diagram commutes because precomposition with the identity is the identity on morphisms:
